---
title: "Session 1"
output: html_document
---

Note: This document is written using Rmarkdown. You can produce HTML and Pdf documents using Rmarkdown and Rstudio. Think documentation, reports, ppt.

## Getting started instructions

- Please install RStudio and R, if you dont have it. [Here](https://www.rstudio.com/products/rstudio/download/)

- Go on Github. Fork my repo and clone that in your computer. 

- The dataset we will work on today is a fun dataset. IMDB Movies. Please go ahead and download the data folder in your project directory. [Download from here](https://drive.google.com/open?id=0B9NCHJOuElCGfllXSFJsd2dFWVdabEdjQVJHSmpKaUVfZUw3VTM1NFMyT0Z6dlIzSVM1MGM)



- Install all the packages

```{r}
# list.of.packages <- c("devtools","readr", "read_xl", "ggplot2", "dplyr", "data.table","rmarkdown", "stringr")
# new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])]
# if(length(new.packages)) install.packages(new.packages)

```

### How to get help when after downloading a package?

- Vignette
- Github


## Read the data

Note:
- Common formats I know about in Biology - Flat files  - tsv, csv, txt, *some random extension of a tab delimited file*, xlsx/xls

- You don't have Databases SQL, JSON etc..

###Two recommended packages 

1. [`readr`](https://github.com/hadley/readr) and [`readxl`](https://github.com/hadley/readxl) for readability and concise
2. [`data.table::fread`](https://github.com/Rdatatable/data.table) for performance and consistency


``` {r}
library(readr)
library(data.table)
titles_df = read_csv("data/titles.csv")
cast_df = fread("data/cast.csv", sep = ",",header = T, verbose = T, stringsAsFactors = F)
rel_dates_df <- read_csv("data/release_dates.csv")
```

That was easy right! 

- Please note the `<-` and `=` assignments. Lets talk! But in practical use cases choose either.
- `<<-` parent scope and `<-` local scope. *what is scope?*



```{r}
str(rel_dates_df)
```

This is the structure command to look at meta deta for your data. We need to understand common data types in R

- Atomic Vectors
  + Boolean/Logical #R is case sensitive T, F, TRUE, FALSE
  + Numeric #double
  + Integer #integer
  + Character
  + Complex! # 2+ 1i
- Matrices and array #I personally did not use them so much
- Lists (trivial and common data strucure)
- Factors
  + levels()
- Dataframe

### Q. How to know what class is the variable. 

```{r}
class(rel_dates_df)
class(rel_dates_df$title)
```

### What are some basic functions to check my variables, structure, dimensions, peek into my data?


```{r}
str(rel_dates_df)
dim(cast_df)
nrow(titles_df)
ncol(rel_dates_df)
names(cast_df)
head(cast_df)
tail(cast_df)
```

This is quick sanity check of the data. 


# Lets Take a break.! [5 mins]

------------------------------------------------------------------
------------------------------------------------------------------

### List all of the "Batman" movies.

```{r}
library(dplyr)
library(stringr)
batman_movies <- filter(titles_df, str_detect(title, "Batman"))
batman_movies
```



### Which is the earliest movie on batman?

```{r}
arrange(batman_movies, year)[1]
```


### Which are the most recent 5 movies on 'batman'?

```{r}
batman_movies <- arrange(batman_movies, desc(year))
slice(batman_movies, 1:5)
```


### How many movies were made from 1990 through 1999?
```{r}
nrow(filter(titles_df, year >= 1990, year <= 1999))
nrow(filter(titles_df, year >= 1990 & year <= 1999))
```



### Which actors were there in 'Jurassic Park'<fill your fav movie> with non null values?

```{r}
jp_actors <- filter(cast_df, title == 'Jurassic Park', type == 'actor', is.na(n))
select(jp_actors, name, character)
```



###Some practice questions for you 
1. How many people have played a role called "The Dude"?
2. How many people have played a role called "The Stranger"?
3. How many roles has Tom Cruise played throughout his career?
4. How many roles were avilable for actresses in the 1950s?
5. How has the rate of movies being released changed over time?
6. What is the ratio of actor to actress in movies?



Lets Take a break! [15 mins]


### What are the ten most common movie names of all time?
```{r}
titles_df %>%
  group_by(title) %>%
  summarise(title_count=n()) %>%
  arrange(desc(title_count)) %>%
  slice(1:10)
```


### What is the median of n for brad pitt and angelina jolly?

```{r}

cast_df %>%
  filter(tolower(name)=='brad pitt' | tolower(name)=='angelina jolie')  %>%
  group_by(name) %>%
  summarise(median_n = median(n)) 
```









### How has the rate of movies being released changed over time in India <fill your country of choice>?



```{r}
library(ggplot2)
qplot(rel_dates_df$year)
```


*Note : Don't start with qplot. You will not be able to learn the grammer properly. Depends on your knowledge*


##Historgram





###Bar Graph


#### Plot the number of films that have been released each decade over the history of cinema.

```{r}
library(ggplot2)
rel_dates_df %>%
  mutate(decade = as.factor(year %/% 10)) %>% # I havd to look at stackoverflow for remainder :P
  group_by(decade) %>%
  summarise(decade_count = n()) %>%
  ggplot(aes(x=decade, y= decade_count)) +
    geom_bar(stat="identity")

```




###Line Graph


#### How do India and USA rank based on number of releases over years?

```{r}

rel_dates_df %>%
  filter(tolower(country)=='india' | tolower(country)=='usa')  %>%
  mutate(country = as.factor(country)) %>%
  group_by(country, year) %>%
  summarise(annual_count = n()) %>%
  ggplot(aes(x = year, y = annual_count)) +
    geom_line(stat="identity") + 
    facet_wrap(~country)
```



### Scatter Plot


#### Plot the n-values of the roles that Scarlett Johansson has played over her career.

```{r}
cast_df %>%
  filter(name=="Scarlett Johansson") %>%
  ggplot(aes(x=year, y=n)) +
  geom_point()
```













