---
title: "Session 1"
output: html_document
---



Note: This is written using Rmarkdown. You can produce HTML and Pdf documents using Rmarkdown and Rstudio. Think documentation, reports, ppt.

The dataset we will work on today is a fun dataset. IMDB Movies. Please go ahead and download the data folder in your project directory. [Download from here](https://drive.google.com/open?id=0B9NCHJOuElCGfllXSFJsd2dFWVdabEdjQVJHSmpKaUVfZUw3VTM1NFMyT0Z6dlIzSVM1MGM)

## Read the data

Common formats I know about in Biology - Flat files  - tsv, csv, txt, <some random extension of a tab delimited file>, xlsx/xls
You don't have Databases SQL, JSON etc..

###Two recommended packages 

1. `readr` and `read_xl` for readability and concise
2. `data.table::fread` for performance and consistency


``` {r}
library(readr)
library(stringr)
titles_df = read_csv("data/titles.csv")
cast_df = read_csv("data/cast.csv")
rel_dates_df <- read_csv("data/release_dates.csv")
```

That was easy right! 

- Please note the `<-` and `=` assignments. Lets talk! But in practical use cases choose either.
- `<<-` parent scope and `<-` local scope



```{r}
str(rel_dates_df)
head(rel_dates_df)
```

This is the structure command to look at meta deta for your data. We need to understand common data types in R

- Atomic Vectors
  - Boolean/Logical
  - Numeric #double
  - Integer #integer
  - Character
  - Complex! # 2+ 1i
  
  
- Matrices and array #I personally did not use them so much
- Lists (trivial and common data strucure)
- Factors
  - levels()
- Dataframe


```{r}





```





### How has the rate of movies being released changed over time?

### What is the ratio of actor to actress in movies?

### List all of the "Batman" movies from earliest to most recent.

### How do countries rank based on number of releases?

### How has the rate of movies being released changed over time in India <fill your country of choice>?

### How many roles were there in '' <

```{r}
library(ggplot2)
qplot(rel_dates_df$year)

```










